---
title: "FRAP_CalMapper"
output: 
  html_document:
    toc: TRUE
date: "2023-09-15"
---

```{r, include = F}
require(sf)
require(tidyverse)
require(mapview)
require(VennDiagram)
```

# Load data
```{r}
load(file = "Rdata/frap_2020_dis_Rx.Rdata")
load(file = "Rdata/CalMapper_activities_fire_2020.Rdata")
load(file = "Rdata/frap_2020_Rx.Rdata")
load(file = "Rdata/NF.Rdata")
```

It's too impossible to actually find overlaps of activity acres so I'm going to stick with dissolving them and finding footprint acreage overlap

## Dissolve 
```{r}
act_20_diss <- act_2020 %>% 
  st_union()
```

# Transform to match
```{r}
act_20_diss <- st_transform(act_20_diss, st_crs(frap_20))
```


## Overlap
```{r}
overlap <- st_intersection(frap_diss_2020, act_20_diss)
overlap_ha <- as.numeric(st_area(overlap)/10000) %>% 
  round()
overlap_ha
```

## Find areas of each
```{r}
frap_20_ha <- as.numeric(st_area(frap_diss_2020)/10000) %>% 
  round()
calmapper_20_ha <- as.numeric(st_area(act_20_diss)/10000) %>% 
  round()
```

```{r}
paste("total area of FRAP dissolved polygons is", frap_20_ha, "ha")
paste("total area of CAL MAPPER dissolved polygons is", calmapper_20_ha, "ha")
```

# Plot with National Forests to make sure it looks right
```{r}
mapview::mapview(list(act_20_diss, NF, frap_diss_2020),
                 col.regions=list("red","blue", "green"),
                 col=list("red","blue", "green"),
                 alpha.regions = 0.4)
```


# Function to make Venn diagram
```{r}
venn_calmap <- function(){
  grid.newpage()
  draw.pairwise.venn(area1 = frap_20_ha,
                   area2 = calmapper_20_ha,
                   cross.area = overlap_ha,
                   #category = c("", ""),
                   fill = c("red", "green"),
                   cex = 1)
}
venn_calmap()
```

```{r}
file <- "Venn_FRAP_CalMAPPER.jpeg"
grid.newpage()
png(file = file, width = 1000, height = 1000)
venn_calmap()
dev.off()
```

# Examine treatments in FRAP that don't overlap
```{r}
frap_only <- st_difference(frap_20, act_20_diss)
```

```{r}
frap_only %>% head()
```

```{r}
mapview::mapview(list(frap_20, frap_only), col.regions=list("red","blue"),col=list("red","blue"))
```

```{r}
frap_only <- frap_only %>% 
  mutate(area_m2 = st_area(frap_only)) %>% 
  arrange(desc(area_m2))
```

```{r}
head(frap_only)
```

```{r}
write.csv(frap_only %>% st_drop_geometry(), file = "FRAP_not_CalMapper.csv", row.names = F)
```

# Make a Venn diagram using treated_acres columns instead of GIS acreage

frap_20 and act_2020

```{r}
frap_only_ID <- frap_20 %>% 
  filter(!TREATMENT_ %in% act_2020$TREATMENT_ID)
frap_only_ID %>% summarize(sum(TREATED_AC))
```

```{r}
overlap_id_ac <- frap_20 %>% 
  st_drop_geometry() %>% 
  filter(TREATMENT_ %in% act_2020$TREATMENT_ID) %>% 
  summarize(sum(TREATED_AC))
overlap_id_ac
```


```{r}
overlap_id_ac <- act_2020 %>% 
  st_drop_geometry() %>% 
  filter(TREATMENT_ID %in% frap_20$TREATMENT_) %>% 
  summarize(sum(TREATED_ACRES))
overlap_id_ac
```


```{r}
frap_id_ac <- frap_20 %>% 
  st_drop_geometry() %>% 
  summarize(sum(TREATED_AC))
frap_id_ac
```

```{r}
calmapper_id_ac <- act_2020 %>% 
  st_drop_geometry() %>% 
  summarize(sum(TREATED_ACRES))
calmapper_id_ac
```

```{r}
venn_calmap_IDs <- function(){
  grid.newpage()
  draw.pairwise.venn(area2 = frap_id_ac,
                   area1 = calmapper_id_ac,
                   cross.area = overlap_id_ac,
                   #category = c("", ""),
                   fill = c( "green", "red"),
                   cex = 0,
                   ext.text = T,
                   euler.d = T,
                   scaled = T)
}
venn_calmap_IDs()
```

```{r}
file <- "Venn_FRAP_CalMAPPER_IDs.jpeg"
grid.newpage()
png(file = file, width = 1000, height = 1000)
venn_calmap_IDs()
dev.off()
```

