---
title: "fire data"
author: "Lisa Rosenthal"
date: "2023-10-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(fs)

```

Load up CA boundary

```{r}
# California boundary
ca_path <- file.path(ref_path, 'CA boundary/ca-state-boundary/CA_State_TIGER2016.shp')
CA <- st_read(ca_path) 
```



Set parameters 

I want fire data from the year of interest (2020). The CRS is WGS 84 / Pseudo-Mercator, from the CA boundary file. Buffering data with a radius equal to half the VIIRS resolution.
```{r}
yr <- 2020
useCRS <- st_crs(CA) #"epsg:3857" 
radius <- 375/2

prep_polys <- function(polys, buffer = T){
  polys <- polys %>% 
    st_transform(useCRS) %>% 
    st_make_valid()
  if(buffer)polys <- st_buffer(polys, radius)
  return(polys)
}
```

## Prep fire data

### FRAP

I want both wildfire and Rx fire data. 
```{r}
# load of data
frap_path <- file.path(ref_path, 'FRAP/fire221gdb/fire22_1.gdb')
st_layers(frap_path)
frap_wf <- st_read(frap_path, 'firep22_1')
# needed to read in rx, write as geojson, reimport. had issues with multisurfaces.
# frap_rx <- st_read(frap_path, 'rxburn22_1')
# st_write(frap_rx, file.path(ref_path, 'FRAP/fire221gdb/rxburn22.geojson'))
frap_rx <- st_read(file.path(ref_path, 'FRAP/fire221gdb/rxburn22.geojson'))

# clean up the data
# filter by date, remove useless columns, buffer by 375/2
frap_rx2020 <- frap_rx %>% 
  filter(year(START_DATE) == yr | year(END_DATE) == yr) %>% 
  select(-(RX_CONSUM:Shape_Area)) %>% 
  prep_polys()
frap_wf2020 <- frap_wf %>% 
  filter(year(ALARM_DATE) == yr | year(CONT_DATE) == yr) %>% 
  select(-(CAUSE:Shape_Area)) %>% 
  prep_polys

# put into a list
fires <- list(frap_rx = frap_rx2020, 
     frap_wf = frap_wf2020)
```

### CalMAPPER

I already worked with CalMAPPER to address question 1 and 3. Save time and load that up. 

```{r}
CM_burn2020_buffered <- read_rds(file.path(ref_path, 'CalMAPPER/burnactivities_2020_buffered.rds'))
CM_burn2020_buffered <- prep_polys(CM_burn2020_buffered, F)

# add to list of fire data
fires$CalMAPPER <- CM_burn2020_buffered

```

### FACTS

The shapefile version of the data isn't complete and I can't download the gdb for some reason. Carmen sent me the data from her pull from a year ago, so using that. 

```{r}

# load data
path <- dir_ls(file.path(ref_path, 'facts'), 
               recurse = T, glob = "*Rdata")
load(path, verbose = T)

# filter more and prep to join list of fire data
facts20 <- facts_1920 %>% 
  filter(year(DATE_COMPL) == 2020, 
         !st_is_empty(.)) %>% 
  select(FACTS_ID, NAME, ACTIVITY_C, ACTIVITY, DATE_COMPL, COUNTY_NAM) %>% 
  mutate(polygon_acres = as.numeric(st_area(.) %>% units::set_units('acres')), 
         .before = geometry)

facts20_prepped <- prep_polys(facts20)

# add to list of fire data
fires$facts <- facts20_prepped


# this is what I'd do.
# # read in data
# facts <- st_read(dir_ls(file.path(ref_path, 'facts'), 
#                 recurse = T, glob = "*Common*shp"))
# 
# # this is huge dataset and I don't need all the rows or fields. reduce it down. 
# # rx fire acivity codes
# codes <- c(1111, 1112, 1113, 1130, 3340, 4471, 4481, 4491, 4541, 7050, 7015,6101, 2540)
# facts_sm <- facts %>% 
#   select(FACTS_ID, NAME, GIS_ACRES, ACTIVITY_C, ACTIVITY, DATE_COMPL, STATE_ABBR) %>% 
#   filter(ACTIVITY_C %in% codes) %>% 
#   filter(!st_is_empty(.), STATE_ABBR == 'CA') %>% 
#   arrange(DATE_COMPL)
# 
# # look at it
# facts_sm
# 
# rm(facts); gc() # remove unused stuff from memory
# 
# # now that its smaller, transform crs and spatially filter to CA
# dim(facts_sm)
# facts_smt <- st_transform(facts_sm, useCRS)
# facts_CA <- facts_smt[CA,]



```


Export the fire data

```{r}
write_rds(fires, 'outputs_spatial/vector/fire_records_2020.rds')
```

