---
title: "prep_PFIRS"
output: html_document
date: "2023-09-05"
---

```{r, include = F}
require(tidyverse)
require(sf)
require(readxl)
require(mapview)
```

# 2019 Data pulled in 2022
```{r}
file <- paste0(ref_path, "/PFIRS/PFIRS 2019-2020 CT pulled 2022.xlsx")
xy_old <- read_excel(file)
```

```{r}
xy_old %>% head()
```

## Fix column names
```{r}
xy_old <- xy_old %>% 
  rename(Burn_Date = `Burn Date`) %>% 
  rename(Burn_Unit = `Burn Unit`) %>% 
  rename(Acres_Burned = `Acres Burned`) %>% 
  rename(Fuel_Type = `Fuel Type`) %>% 
  rename(Total_Tons = `Total Tons`) %>% 
  rename(Burn_Type = `Burn Type`) 
```

## Make lat and long numeric and remove unk rows
```{r}
xy_old <- xy_old %>% 
  filter(!Longitude == "UNK") %>% 
  mutate(Longitude = as.numeric(Longitude)) %>% 
  mutate(Latitude = as.numeric(Latitude))
```


## Fix dates that should be negative
```{r}
xy_old <- xy_old %>% 
  mutate(Longitude = ifelse(Longitude > 0, Longitude*(-1), Longitude))
```

## Add year column
```{r}
xy_old <- xy_old %>% 
  mutate(Year = year(Burn_Date))
```



## Filter to 2019
```{r}
xy_old_2019 <- xy_old %>% 
  filter(Year == 2019)
```

```{r}
xy_old_2019 %>% head()
```


# Delete duplicates

From Jason Branz: "Typically if there are multiple records with the same date, burn name, and acres, it’s a duplicate." 

## Look at duplicates
```{r}
xy_old_2019 %>% 
  select(Burn_Date, Burn_Unit, Acres_Burned) %>% 
  filter(duplicated(.))
```

## Remove duplicates
```{r}
xy_old_2019 <- xy_old_2019 %>% 
  distinct(Burn_Date, Burn_Unit, Acres_Burned, .keep_all = T)
```



## Make spatial
```{r}
sf_old_2019 <- st_as_sf(xy_old_2019, coords = c("Longitude", "Latitude"), crs = 4326)
```


## Export to shapefile
```{r}
st_write(sf_old_2019, dsn = "shapefiles", layer = "PFIRS_2019_pulled2022", driver = "ESRI Shapefile", delete_layer = T)
```

# 2019 Data pulled in 2023
```{r}
xy_new <- read_excel(paste0(ref_path, "/PFIRS/PFIRS 2017-2022 pulled 2023.xlsx"))
```

```{r}
xy_new
```

## Fix column names
```{r}
xy_new <- xy_new %>% 
  rename(Burn_Date = `Burn Date`) %>% 
  rename(Burn_Unit = `Burn Unit`) %>% 
  rename(Acres_Burned = `Acres Burned`) %>% 
  rename(Fuel_Type = `Fuel Type`) %>% 
  rename(Total_Tons = `Total Tons`) %>% 
  rename(Burn_Type = `Burn Type`) 
```


## Fix dates that should be negative
```{r}
xy_new <- xy_new %>% 
  mutate(Longitude = ifelse(Longitude > 0, Longitude*(-1), Longitude))
```

## Add year column
```{r}
xy_new <- xy_new %>% 
  mutate(Year = year(Burn_Date))
```


## Remove duplicates
```{r}
xy_new <- xy_new %>% 
  distinct(Burn_Date, Burn_Unit, Acres_Burned, .keep_all = T)
```

## Make spatial and save

### 2019

#### Filter to 2019
```{r}
xy_new_2019 <- xy_new %>% 
  filter(Year == 2019)
```


```{r}
xy_new_2019 %>% head()
```

#### Make spatial
```{r}
sf_new_2019 <- st_as_sf(xy_new_2019, coords = c("Longitude", "Latitude"), crs = 4326)
```

#### Export to shapefile
```{r, warning=F}
st_write(sf_new_2019, "shapefiles", "PFIRS_2019_pulled2023", driver = "ESRI Shapefile", delete_layer = T)
```

### All years

#### Make spatial
```{r}
sf_new <- st_as_sf(xy_new, coords = c("Longitude", "Latitude"), crs = 4326)
```

#### Export to shapefile
```{r, warning=F}
st_write(sf_new, "shapefiles", "PFIRS_2017-2022_pulled2023", driver = "ESRI Shapefile", delete_layer = T)
```

#### Save as Rdata
```{r}
save(sf_new, file = "PFIRS_2017-2022_pulled2023.Rdata")
```

# Mask out federal lands

## Crop to state of CA
```{r}
CA <- st_read(dsn = paste0(ref_path, "/CA boundary/ca-state-boundary/", layer = "CA_State_TIGER2016.shp"))
CA <- st_transform(CA, st_crs(sf_new))
```

```{r}
sf_new <- st_intersection(sf_new, CA)
```


```{r}
load(file = "Rdata/NF.Rdata")
```

```{r}
NF <- st_transform(NF, st_crs(sf_new))
```

```{r}
sf_use_s2(FALSE)
```

```{r}
sf_new_noNF <- st_difference(sf_new, NF)
```

```{r}
mapview::mapview(list(NF, sf_new_noNF), col.regions=list("red","blue"),col=list("red","blue"))
```

#### Save as Rdata
```{r}
save(sf_new_noNF, file = "Rdata/PFIRS_2017-2022_pulled2023_noNF.Rdata")
```


