---
title: "VIIRS prep and overview"
author: "Lisa Rosenthal"
date: "2023-09-13"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(terra)
library(fs)
library(tidyverse)
```

# Goal

-   Check out how VIIRS data in 2020 is spatially distributed across CA
-   Mask out crop areas in order to compare to CalMAPPER, etc.
-   See if there's an easy way to detect "persistent points" as a way to potentially limit false fire detections

# Import Data

```{r}

# VIIRS fire detects
viirs_dir <- file.path(ref_path, 'VIIRS/CA_2020')
viirs_shps <- dir_ls(viirs_dir, recurse = T, glob = '*.shp')
viirs_list <- map(viirs_shps, st_read) %>% 
  set_names(c('NOAA', 'Suomi'))

# crop layer
cdl_path <- file.path(ref_path, "NASS/CDL/2020_30m_cdls") %>% dir_ls(glob = '*.tif')
cdl <- rast(cdl_path)

# California boundary
ca_path <- file.path(ref_path, 'CA boundary/ca-state-boundary/CA_State_TIGER2016.shp')
CA <- st_read(ca_path)

```

# Crop to CA and mask out agriculture

Just include VIIRS and crop data to CA. Remove VIIRS points in agricultural areas according to the 2020 CDL dataset. I omitted crops (except Xmas tree farms), development, and water. Keeping wildlands and "pasture/grass".

```{r}
# transform vector layers to raster crs
CA_t <- st_transform(CA, crs(cdl))
viirs_list_t <- map(viirs_list, ~ st_transform(.x, crs(cdl)))

# crop VIIRS to CA boundary
viirs_list_CA <- map(viirs_list_t, ~ .x[CA_t, ])

# crop CDL to CA boundary
cdl_CA <- crop(cdl, CA_t)

# create a crop mask
levels_table <- levels(cdl_CA)[[1]]
head(levels_table)
# levels_table$Class_Names

# make a vector containing the classes you want to keep
nonAgVeg <- c('Pasture/Grass', 'Forest', "Shrubland", "Barren", "Christmas Trees",
              "Nonag/Undefined", "Deciduous Forest", "Evergreen Forest", "Mixed Forest",
              "Grassland/Pasture", "Woody Wetlands", "Herbaceous Wetlands"
              )
nonAgVeg_values <- levels_table$value[levels_table$Class_Names %in% nonAgVeg]

# reclassify the raster to be non-ag vs. not
mat1 <- cbind(nonAgVeg_values, 1)
cdl_CA_mask <- classify(cdl_CA, mat1, others = 0)
plot(cdl_CA_mask)

# remove VIIRS points outside the non-ag area
f_extract <- function(mask, points){
  vals <- extract(mask, points)
  nonAg_points <- points[vals$Class_Names == 1,]
  return(nonAg_points)
}
viirs_list_nonAg <- map2(cdl_CA_mask, viirs_list_CA, f_extract) %>% 
  set_names(names(viirs_list_CA))
```

# clean up VIIRS data

Fix time variables (UTC instead of local), merge satellites

```{r}
# let's merge Suomi and NOAA-20 satellites
viirs_nonAg <- bind_rows(viirs_list_nonAg)

# transform back into original crs too
viirs_nonAg <- st_transform(viirs_nonAg, st_crs(viirs_list$NOAA))

# get datetime in PST. time / date currently in UTC.
viirs_nonAg <- viirs_nonAg %>%
  mutate(
    # get datetime in PST
    .before = ACQ_DATE,
    hour = str_sub(ACQ_TIME, 1, 2),
    minute = str_sub(ACQ_TIME, 3, 4),
    time = paste(hour, minute, sep = ':'),
    datetime = ymd_hm(paste(ACQ_DATE, time)) %>% with_tz('US/Pacific'),
    date = date(datetime),
    # detects in the middle of the night (e.g. 3am) should be considered night before
    date_adj = case_when(DAYNIGHT == 'N' ~ date - 1, T ~ date),
    jday = yday(date),
    jday_adj = ifelse(DAYNIGHT == 'N', jday - 1, jday),
    jday_daynight_adj = paste(jday_adj, DAYNIGHT, sep = '_')
  ) %>%
  select(-c(hour, minute, time, ACQ_DATE, ACQ_TIME)) 

# dummy check that the time makes sense
viirs_nonAg %>% 
  as_tibble %>% 
  sample_n(4000) %>% 
  mutate(hour = hour(datetime)) %>% 
  ggplot(aes(DAYNIGHT, hour)) +
  geom_jitter() +
  labs(title = 'Dummy checking time conversion', 
       subtitle = '**PST hour is consistent with DAYNIGHT column')

```

# Check out the spatial distribution of VIIRS data

For a quick look, let's look at just one week...Aug1-15 2020.

```{r}

viirs_filtered <- viirs_nonAg %>% 
  filter(month(datetime) == 8, day(datetime) <= 15)
mapview::mapview(viirs_filtered, zcol = 'jday', cex = 3)

```

# Export data

```{r}

# viirs--CA, non-Ag
write_rds(viirs_nonAg, file = file.path(viirs_dir, 'viirs_nonAg.rds'))

# crop mask
writeRaster(cdl_CA_mask, file.path(ref_path, "NASS/CDL/cropmask_2020.tif"))

```
