---
title: "CalMAPPER"
output: 
  html_document:
    toc: TRUE
date: '2022-07-26'
editor_options: 
  chunk_output_type: console
---

```{r, include = F}
require(tidyverse)
require(lubridate)
require(sf)
require(fs)
require(mapview)
```

# Description

This script merges CalMAPPER activity data with treatment polygons. This is a necessary step before analyzing prescribed fire data in CalMAPPER. 


# Define path to Calmapper and to the gdb file
```{r}
calmapper_dir <- fs::dir_ls(ref_path, recurse = T, glob = '*CalMAPPER', type = 'directory')
calmapper_gdb <- fs::dir_ls(calmapper_dir, recurse = T, glob = '*.gdb')
if(length(calmapper_gdb) > 1){
  stop( "There's more than one CalMAPPER .gdb file. script assumes only 1.")
}

```

```{r}
st_layers(calmapper_gdb)
act <- st_read(calmapper_gdb, layer = 'CMDash_Activities')
trt <- st_read(calmapper_gdb, layer = 'CMDash_TreatmentPols')
```


# Clean
## Remove non-treatment "activities"
```{r}
act <- act %>% 
  filter(!ACTIVITY_DESCRIPTION %in% c("GIS Validation", "Education Outreach", "Project Administration", "Planning Meeting", "Public Contacts", "Water Site Development"))
```

# Merge Rx fire Activity data with treatment polygons
```{r}
act_sf <- right_join(trt %>% select(PROJECT_ID, TREATMENT_ID), 
           act)
```

# Filter to fire 

## Find relevant treatment names
```{r}
act_sf %>% 
  st_drop_geometry() %>% 
  group_by(ACTIVITY_DESCRIPTION) %>% 
  count() %>% 
  print(n=50)
```

## Filter by treatment
```{r}
act_sf_fire <- 
  act_sf %>% 
  filter(ACTIVITY_DESCRIPTION %in% c("Broadcast Burn", "Cultural Burning", "Pile Burning"))
```

# Create DURATION and YEAR columns

## DURATION
```{r}
act_sf_fire <- act_sf_fire %>% 
  mutate(DURATION = as.Date(ACTIVITY_END) - as.Date(ACTIVITY_START)+1)
```

## Add YEAR column
```{r}
act_sf_fire <- act_sf_fire %>% 
  mutate(YEAR = year(ACTIVITY_END))
```

## Check
```{r}
act_sf_fire %>% 
  select(ACTIVITY_START, ACTIVITY_END, DURATION, YEAR) %>% 
  head() %>% 
  nrow()
```


# Save
```{r}
save(act_sf_fire, file = "Rdata/CalMapper_activities_fire.Rdata")
```

```{r}
write.csv(act_sf_fire %>% st_drop_geometry(), file = "~/Reference data/CALFIRE spatial data/CalMapper/activities_fire.csv", row.names = F)
write.csv(trt %>% st_drop_geometry(), file = "~/Reference data/CALFIRE spatial data/CalMapper/treatments_fire.csv", row.names = F)
```

# Are there Rx fires recorded in Treatments data but not in Activities data or vice versa?
```{r}
act_broadcast <- act_sf_fire %>% 
  filter(ACTIVITY_DESCRIPTION == "Broadcast Burn")
```


```{r}
trt %>% 
  st_drop_geometry() %>% 
  group_by(TREATMENT_OBJECTIVE) %>% 
  count()
```


```{r}
trt_broadcast_complete <- trt %>% 
  filter(TREATMENT_OBJECTIVE == "Broadcast Burn") %>% 
  filter(ACTIVITY_STATUS == "Complete")
```

## Check if there are treatment IDs in `act_broadcast` that aren't in `trt` -- this shows whether the merge was complete 
```{r}
act_broadcast %>% 
  filter(!TREATMENT_ID %in% trt$TREATMENT_ID) %>% 
  nrow()
```

```{r}
act_broadcast %>% 
  filter(!TREATMENT_NAME %in% trt$TREATMENT_NAME) %>% 
  nrow()
```


## Check if there are treatment IDs in completed broadcast treatments that aren't in activities
```{r}
trt_broadcast_complete %>% 
  filter(!TREATMENT_ID %in% act_broadcast$TREATMENT_ID) %>% 
  mapview()
```

```{r}
trt_broadcast_complete %>% 
  filter(!TREATMENT_NAME %in% act_broadcast$TREATMENT_NAME) %>% 
  nrow()
```

## Export those to put in .ppt
```{r}
missing_from_activities <- trt_broadcast_complete %>% 
  filter(!TREATMENT_ID %in% act_broadcast$TREATMENT_ID) %>% 
  select(CALMAPPER_ID, PROJECT_NAME, TREATMENT_NAME, TREATMENT_OBJECTIVE, PROJECT_STATUS, ACTIVITY_STATUS, PROJECT_START_DATE, PROJECT_END_DATE, TREATMENTAREA_ACRES) %>% 
  st_drop_geometry()
missing_from_activities %>% 
  summarize(missing_acres = sum(TREATMENTAREA_ACRES))
write_excel_csv(missing_from_activities, "calmapper_trt_not_act.xls")
```

## Visualize broadcast treatments with no overlapping activities
```{r}
map <- mapview(list(trt_broadcast_complete, act_broadcast), col.regions=list("red","blue"),col=list("red","blue"))
map
```

It's safe to use the Activities information and not the Treatment information.

# Write activities data to gdb
```{r}
st_write(act_sf_fire, dsn = calmapper_gdb, layer = "CalMapper_fire_act", delete_layer = T)
```

# Save 2020 to use for Venn Diagram
```{r}
act_2020 <- act_sf_fire %>% 
  filter(YEAR == 2020)
```

# Mask out federal lands
```{r}
dsn = paste(ref_path, "/National Forest Boundaries", sep = "")
NF <- st_read(dsn = dsn, "NF_CA_diss")
NF <- st_transform(NF, st_crs(act_2020))
save(NF,  file = "Rdata/NF.Rdata")
```

```{r}
act_2020 <- st_difference(act_2020, NF)
```

# Save
```{r}
save(act_2020, file = "Rdata/CalMapper_activities_fire_2020.Rdata")
```

