---
title: "Q2_VIIRS_wo_records"
author: "Lisa Rosenthal"
date: "2023-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(fs)
library(terra)
library(mapview)
library(fasterize)

```

### Functions
```{r}

# manipulations dont like it when geometries aren't valid. and they don't work with NA validity.
make_valid <- function(sf_object) {
  validity <- st_is_valid(sf_object)
  if (sum(is.na(validity)) > 0) {
    stop("can't proceed. Contains rows where validity is NA.")
  } else{
    if (all(validity)) {
      return(sf_object)
    } else{
      sf_object <- st_make_valid(sf_object)
      print(glue::glue("there were {sum(validity == F)} not valid geometries"))
      return(sf_object)
    }
  }
}

# transform to CA crs, select fields you want, make valid
prep_polys <- function(polygon, columns, points = CA){
  message("transforming polygons")
  polygon <- polygon %>% 
  st_transform(st_crs(points)) %>% 
  select(all_of(columns))
  
  message("making polygons valid")
  make_valid(polygon)
}


# extract data from vector layers
extract_polys <- function(polygons, points) {
  #points = viirs_sample
  #polygons = vector_layers$powerplants
  
  intersects <- st_intersects(points, polygons, sparse = T)
  if ( length(unlist(intersects)) == 0)
    return(warning("No intersections"))
  else{
    intersections_list <- map(intersects, unlist)
    
    # error if there's more than 1 polygon intersecting a point
    if(length(unlist(intersections_list)) > nrow(points)){
     warning("Point(s) overlapped with more than 1 polygon. By default, choosing first one.")
    # make sure points with no intersection become NA
      }else{
      polygonIDs <- map_vec(intersections_list, 
                           ~ ifelse(sum(.x) == 0, NA, .x[1]) )
      newcols <- st_drop_geometry(polygons)[polygonIDs, ]
      return(as_tibble(newcols))
    }
  }
}

```
## Goal

Create a cleaned up version of VIIRS points consisting of high confidence fire points. 

Remove points that are
- On ag land, urban, water (from NASS CDL)
- campgrounds
- power plants
- solar farms

Other layers that could be important
- Jurisdictions/Land management agency


to rid stationary hotspots, could rasterize the points and calculate the density of viirs points during the non-rx fire months. >4 unique days get thrown out? need to mask out wildfire.

## Load CA boundary data

Reminder that I need to redownload VIIRS to completely cover the southern part of CA!

```{r}
# California boundary
ca_path <- file.path(ref_path, 'CA boundary/ca-state-boundary/CA_State_TIGER2016.shp')
CA <- st_read(ca_path) 

# VIIRS fire detects
viirs_dir <- file.path(ref_path, 'VIIRS/CA_2020')
viirs_shps <- dir_ls(viirs_dir, recurse = T, glob = '*.shp')
viirs_shps <- viirs_shps[!grepl("viirs_nonAg", viirs_shps)] # don't want that file
viirs <- map_df(viirs_shps, st_read) %>% st_transform(st_crs(CA))
viirs <- viirs[CA,]

# get viirs into PST, not UTC
viirs <- viirs %>%
  mutate(
    # get datetime in PST
    .before = ACQ_DATE,
    hour = str_sub(ACQ_TIME, 1, 2),
    minute = str_sub(ACQ_TIME, 3, 4),
    time = paste(hour, minute, sep = ':'),
    datetime = ymd_hm(paste(ACQ_DATE, time)) %>% with_tz('US/Pacific'),
    date = date(datetime),
    # detects in the middle of the night (e.g. 3am) should be considered night before
    date_adj = case_when(DAYNIGHT == 'N' ~ date - 1, T ~ date),
    jday = yday(date),
    jday_adj = ifelse(DAYNIGHT == 'N', yday(date_adj), jday),
  ) %>%
  select(-c(hour, minute, time, ACQ_DATE, ACQ_TIME)) 


# vis extent of the two. 
st_make_grid(st_bbox(CA), n = 1) %>% plot(col = 'maroon4')
st_make_grid(st_bbox(viirs), n = 1) %>% plot(add = T, col = 'tan')
plot(st_geometry(CA), add = T)


```

Import the data to extract with the VIIRS points
```{r}

vector_layers <- read_rds(file.path(ref_path, 'merged data layers/vector', 'extract_viirs.rds'))
rasts <- rast(file.path(ref_path, 'merged data layers/raster', 'extract_viirs.tif'))

```

Extract data with VIIRS points. 

```{r}

# raster data
extracted_rasters <- terra::extract(x = rasts, viirs)


# vector data
vector_names <- list("county",
                     "camping",
                     c("power_source", 'power_capacity'),
                     "solar")
tictoc::tic()
extracted_vectors <- map(vector_layers, ~ extract_polys(.x, viirs)) %>% 
  map2(., vector_names, ~ set_names(.x, .y)) %>% # rename the columns
  bind_cols()
tictoc::toc() # 21sec

# Check out how often viirs intersects with these things...
# not much. The points (e.g. camping) probably underesimated, but oh well. 
apply(extracted_vectors, 2, function(x) 100 * mean(!is.na(x)))

viirs_ex <- tibble(viirs, extracted_rasters, extracted_vectors)


```


## Fire data


```{r}
yr <- 2020


frap_path <- file.path(ref_path, 'FRAP/fire221gdb/fire22_1.gdb')
st_layers(frap_path)
frap_rx <- st_read(frap_path, 'rxburn22_1')
frap_wf <- st_read(frap_path, 'firep22_1')

frap_rx2020 <- frap_rx %>% 
  filter(year(START_DATE) == yr | year(END_DATE) == yr)
frap_wf2020 <- frap_wf %>% 
  filter(year(ALARM_DATE) == yr | year(CONT_DATE) == yr)

```


Identify areas with VIIRS hotspots? Make sure they're outside of wildfire perimeters.

To rid stationary hotspots, could rasterize the points and calculate the density of viirs points during the non-rx fire months. >4 unique days get thrown out? It's probably just picking up wildfire, which will get masked out anyways. 

```{r}
months_noRxfire <- c(7:9)
viirs_summer <- viirs %>% 
  filter(month(date_adj) %in% months_noRxfire)
viirs_density <- rasterize(viirs_summer, rasts$SRA, fun = 'length')
plot(viirs_density>= 4)

```



