---
title: "Q2_VIIRS_wo_records"
author: "Lisa Rosenthal"
date: "2023-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(fs)
library(terra)
library(mapview)
library(fasterize)

```

### Functions

```{r}

# extract data from vector layers
extract_polys <- function(polygons, points) {
  # polygons = vector_layers$solar
  # points = viirs[1:5000,"point_id"]
  
  intersects <- st_intersects(points, polygons, sparse = T)
  if ( length(unlist(intersects)) == 0)
    return(warning("No intersections"))
  else{
    intersections_list <- map(intersects, unlist)
    
    # warning if there's more than 1 polygon intersecting a point
    if(any(map_vec(intersections_list, length) > 1)){
      warning("Point(s) overlapped with more than 1 polygon. By default, choosing first one.")}
    # make sure points with no intersection become NA
    polygonIDs <- map_vec(intersections_list, 
                          ~ ifelse(sum(.x) == 0, NA, .x[1]) )
    newcols <- st_drop_geometry(polygons)[polygonIDs, ]
    return(as_tibble(newcols))
  }
}

# add geometry id
add_geomID <- function(sf_object){
  # ensure geometry column is named properly
  st_geometry(sf_object) <- 'geometry'
  # add geom id
  sf_object %>%
  mutate(geom_id = row_number(), .before = geometry) %>% 
  group_by(geometry) %>%
  mutate(geom_id = ifelse(n() > 1, geom_id[1], geom_id)) %>% 
  ungroup()
}
```

## Goal

Create a cleaned up version of VIIRS points consisting of high
confidence fire points.

Remove points that are - On ag land, urban, water (from NASS CDL) -
campgrounds - power plants - solar farms

Other layers that could be important - Jurisdictions/Land management
agency

to rid stationary hotspots, could rasterize the points and calculate the
density of viirs points during the non-rx fire months. \>4 unique days
get thrown out? need to mask out wildfire.

## Load CA boundary data

Reminder that I need to redownload VIIRS to completely cover the
southern part of CA!

```{r}
# California boundary
ca_path <- file.path(ref_path, 'CA boundary/ca-state-boundary/CA_State_TIGER2016.shp')
CA <- st_read(ca_path) 

# VIIRS fire detects
viirs_dir <- file.path(ref_path, 'VIIRS/CA_2020')
viirs_shps <- dir_ls(viirs_dir, recurse = T, glob = '*.shp')
viirs_shps <- viirs_shps[!grepl("viirs_nonAg", viirs_shps)] # don't want that file
viirs <- map_df(viirs_shps, st_read) %>% st_transform(st_crs(CA))
viirs <- viirs[CA,]

# get viirs into PST, not UTC
viirs <- viirs %>%
  mutate(
    # get datetime in PST
    .before = ACQ_DATE,
    hour = str_sub(ACQ_TIME, 1, 2),
    minute = str_sub(ACQ_TIME, 3, 4),
    time = paste(hour, minute, sep = ':'),
    datetime = ymd_hm(paste(ACQ_DATE, time)) %>% with_tz('US/Pacific'),
    date = date(datetime),
    # detects in the middle of the night (e.g. 3am) should be considered night before
    date_adj = case_when(DAYNIGHT == 'N' ~ date - 1, T ~ date),
    jday = yday(date),
    jday_adj = ifelse(DAYNIGHT == 'N', yday(date_adj), jday),
  ) %>%
  select(-c(hour, minute, time, ACQ_DATE, ACQ_TIME)) %>% 
  mutate(point_id = row_number(), .before = LATITUDE)


# vis extent of the two. 
st_make_grid(st_bbox(CA), n = 1) %>% plot(col = 'maroon4')
st_make_grid(st_bbox(viirs), n = 1) %>% plot(add = T, col = 'tan')
plot(st_geometry(CA), add = T)


```

## Extract things around the VIIRS points

Import the data to extract with the VIIRS points

```{r}

vector_layers <- read_rds('outputs_spatial/vector/extract_viirs.rds')
rasts <- rast('outputs_spatial/raster/extract_viirs.tif')

```

Extract data with VIIRS points.

```{r}

# raster data
extracted_rasters <- terra::extract(x = rasts, viirs)



# vector data
vector_names <- list(c("county"),
                     c("camping"),
                     c("power_source", 'power_capacity'),
                     c("solar"))
tictoc::tic()
extracted_vectors <- map(vector_layers, ~ extract_polys(.x, viirs)) %>% 
  map2(., vector_names, ~ set_names(.x, .y)) %>% 
  bind_cols
tictoc::toc() # 26.6sec


# Check out how often viirs intersects with these things...
# not much. The points (e.g. camping) probably underesimated, but oh well. 
apply(extracted_vectors, 2, function(x) 100 * mean(!is.na(x)))

viirs_ex <- tibble(viirs, extracted_rasters, extracted_vectors)


```

## Fire data

I combined fire data in the script `prep_fire_polygons.Rmd`. It combines
FRAP (rx and wildfire), CalMAPPER, and FACTS. Also extract fire data,
similar to above.

Originally I wanted information about all polygons, even those that were
overlapping each other, but I needed to create a loop that took forever.
So, still extracting information on the first intersecting polygon.

```{r}
fires <- read_rds(here::here('outputs_spatial/vector/fire_records_2020.rds'))

# assign a polygon ID to each polygon. Seems like the datasets don't always have a unique ID for each row. 
fires <- map(fires, add_geomID)

# choose the columns to be extracted from each layer
cols_to_keep <- list(
  frap_rx = c("geom_id", "TREATMENT_ID", "TREATMENT_NAME", "AGENCY", "START_DATE", "END_DATE", 'TREATED_AC'),
  frap_wf = c("geom_id", "INC_NUM", "FIRE_NAME", "AGENCY", "ALARM_DATE", "CONT_DATE"),
  CalMAPPER = c("geom_id", "TREATMENT_ID", "AQ_ID", 'pile_burn', 'broadcast_burn', "ACTIVITY_START", "ACTIVITY_END", "duration"), 
  FACTS = c("geom_id", "FACTS_ID", "NAME", "ACTIVITY", "DATE_COMPL")
)
fires <- map2(fires, cols_to_keep, ~ select(.x, all_of(.y)))


# extract fire data. 
tictoc::tic()
extracted_fires <- map(fires, ~ extract_polys(.x, viirs))
tictoc::toc() # 26.6sec

# add prefixes to column names and join.
prefixes <- c("FRAPrx", "FRAPwf", "CM", "FACTS")
add_prefix <- function(df, prefix){
  names(df) <- glue::glue("{prefix}_{names(df)}")
  return(df)
}
extracted_fires <- map2(extracted_fires, prefixes, add_prefix) %>% 
  bind_cols()

```

Ok, now I've extracted info on fires and things that might affect viirs
detections. Let's make sure the extractions make sense adn then merge it
all together.

```{r}
# same number of rows. good. merge it together. 
dim(viirs_ex)
dim(extracted_fires)
viirs_ex <- tibble(viirs_ex, extracted_fires)
```

Filter out VIIRS points that spatially AND temporally overlap with
wildfires for now. Wow, there are still \~49k VIIRS detections.

It looks like there are still points that lay right outside the
wildfires. Remove those too.

```{r}

# these viirs points spatially and temporally overlap with wildfire. Remove for now. 
viirs_wf <- viirs_ex %>% 
  filter(!is.na(FRAPwf_geom_id) &
           date_adj >= FRAPwf_ALARM_DATE & date_adj <= FRAPwf_CONT_DATE) 

# these should theoretically be from things other than wildfire. 
viirs_notwf <- viirs_ex %>% 
  filter(!point_id %in% viirs_wf$point_id) %>% 
   mutate(FRAP = !is.na(FRAPrx_geom_id),
            CM = !is.na(CM_geom_id), 
            FACTS = !is.na(FACTS_geom_id)) %>% 
  mutate(anyPolygons = if_any(FRAP:FACTS, ~ .x == T))
nrow(viirs_notwf)

# visualize points. make sure that they were removed from wildfire. Wow, looks like there are still poinst outside the wildfires. take those out. 
# mapview(vector_layers$counties)
sierraNevada_co <- c("Alpine", "El Dorado", "Nevada", "Placer", "Sierra", "Yuba", "Amador", "Calaveras", "Tuolumne", "Mariposa", "Madera", "Mono", "Fresno", "Tulare", "Inyo", "Plumas", "Butte") 
viirs_notwf %>% 
  filter(county %in% sierraNevada_co) %>% 
  st_as_sf(sf_column_name = "geometry") %>% 
  mapview(zcol = "anyPolygons") +
  mapview(fires$frap_wf)

# buffer wildfire perimeters by 5km and remove those happening during the time of the fire
wf_buffered <- st_buffer(fires$frap_wf, dist = 10000)

likely_wf_points <- st_intersection(
  st_as_sf(viirs_notwf, sf_column_name = "geometry"), 
  wf_buffered) %>% 
  filter(date_adj >= ALARM_DATE & (date_adj-5) <= CONT_DATE) # made the date window 5 day lenient on the backend since I think FRAP sometimes cuts the end date short
viirs_notwf <- viirs_notwf %>% 
  mutate(likely_wf = ifelse(point_id %in% likely_wf_points$point_id, T, F))

# visualize again. looks a bit better
viirs_notwf %>% 
  filter(county %in% sierraNevada_co) %>% 
  st_as_sf(sf_column_name = "geometry") %>% 
  mapview(zcol = "likely_wf") +
  mapview(fires$frap_wf)

# append the wf dataset, remove those from no wf dataset
viirs_wf <- bind_rows(viirs_wf, 
          viirs_ex %>% 
  filter(point_id %in% viirs_notwf$point_id[viirs_notwf$likely_wf]))
viirs_notwf <- filter(viirs_notwf, likely_wf == F)

```

Identify areas with VIIRS hotspots? Make sure they're outside of
wildfire perimeters.

To rid stationary hotspots, could rasterize the points and calculate the
density of viirs points during the non-rx fire months. \>4 unique days
get thrown out? It's probably just picking up wildfire, which will get
masked out anyways.

```{r}
months_noRxfire <- c(7:9)
viirs_summer <- viirs_notwf %>% 
  filter(month(date_adj) %in% months_noRxfire) %>% 
  st_as_sf(sf_column_name = "geometry")
viirs_density <- rasterize(viirs_summer, rasts$SRA, fun = 'length')

# map that density raster to the viirs points
val <- terra::extract(viirs_density, st_as_sf(viirs_notwf, sf_column_name = "geometry"), ID = F)
val <- replace_na(val[,1], 0) 
viirs_notwf <- mutate(viirs_notwf, summer_dens = val)

```

## Compare VIIRS points: those in and out of recorded polygons

How many were in wildfires? In the rx fire polygons? how does this vary
by source? How do these line up with some of the factors that might
cause false detections?

I peaked at a map of the data and it looks like there are hotspots of
detections during the summer months. Not sure the optimal threshold, but
there was a paper that used 4 detections outside the burning months. I
found hotspots around powerplants, but also some areas not mapped, like
landfills, quarries, cement factories. Also, it would be nice if the
data were polygons instaed of points.

Links to additional data:

- [DMR All
Mines](https://gis.data.ca.gov/datasets/cadoc::dmr-all-mines/explore?location=37.321295%2C-122.099462%2C14.33)
- [SWIS (Solid Waste Infrastructure
System)](https://gis.data.ca.gov/datasets/2e8d23425c18401b80f198618e93bffd/explore?location=37.322800%2C-122.085933%2C13.20)
- [WESTCARB GIS DATA](http://gif.berkeley.edu/westcarb/gis-data/)--might have polygons?? Not sure how this data was made or when. 

```{r}
# in wildfire
points_wf <- viirs_wf$point_id
paste("percentage of points associated with wildfire:", round(length(points_wf) / nrow(viirs) * 100, 2)) # 92%!
paste("Number of points NOT associated with wildfire:", nrow(viirs_notwf), "out of", nrow(viirs)) # 38149 out of 477303

# For points outside of wildfire, how many in each polygon type? 
viirs_notwf %>% 
  pivot_longer(FRAP:anyPolygons) %>% 
  select(name, value) %>%  
  group_by(name) %>% 
  summarise(percentage = round(100 *sum(value)/nrow(viirs_notwf), 2))


# responsibility areas, non-ag, camping, power_source, summer detection density 
viirs_notwf_sm <- viirs_notwf %>% 
  mutate(across(c(camping, power_source, solar), ~as.numeric(!is.na(.x))), 
         ag_urban_water = ifelse(non_ag == 0, 1, 0)) %>% 
  select(date_adj, FRP, CONFIDENCE, TYPE, SRA, county, non_ag, ag_urban_water, camping, power_source, solar, summer_dens, FRAP:anyPolygons, geometry)
# print(viirs_notwf_sm, width = Inf)

viirs_notwf_sm %>% 
  group_by(anyPolygons) %>% 
  summarise(n_points = n(),
            perc_points = 100 * n_points/nrow(.),
            across(non_ag:solar, ~ sum(.x)), 
            dens_qlow = quantile(summer_dens, c(.05)),
            dens_qhi = quantile(summer_dens, c(.95))
            )


# dig into summer density a little more. I'm not sure what a good threshold is
global_labels <- labeller(
  anyPolygons = c("FALSE" = "no Rx record", "TRUE" = "Rx fire record"), 
  power_source = c("0" = "no powerplant", "1" = "yes powerplant"),
  ag_urban_water = c("0" = "natural", "1" = "ag, urban, or water")
)

ggplot(viirs_notwf_sm, aes(summer_dens)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~anyPolygons, labeller = global_labels)
ggplot(viirs_notwf_sm, aes(summer_dens)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~ag_urban_water, labeller = global_labels)
# powerplants have more persistent VIIRS points, but still plenty of spots with unexplained hotspots. 
ggplot(viirs_notwf_sm, aes(summer_dens)) +
  geom_histogram(binwidth = 1) +
  facet_grid(~power_source, labeller = global_labels)
 
# how about the quality fields taht VIIRS provides?
viirs_notwf_sm %>% 
  filter(!is.na(TYPE)) %>% 
  ggplot(aes(summer_dens)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~TYPE)

# does type line up with powerplants? 
viirs_notwf_sm %>% 
  filter(!is.na(TYPE)) %>% 
  count(TYPE, power_source) %>% 
  arrange(power_source)


# # map it? 
# mapview(st_as_sf(viirs_notwf_sm, sf_column_name = "geometry"),
#         zcol = "summer_dens", cex = 2) +
# mapview(st_as_sf(viirs_notwf_sm, sf_column_name = "geometry"),
#         zcol = "TYPE", cex = 2) +
# mapview(vector_layers$powerplants)
#  
```

Lets remove points that coincide with possible false positive sources. Now map it and explore.

```{r}
# remove ag/urban/water, solar, power plants, summer dens > 4. 
#print(viirs_notwf, width = Inf)

# points
viirs_notwf_f <- viirs_notwf %>% 
  filter(non_ag == 1, is.na(camping), is.na(power_source), is.na(solar), 
         summer_dens < 4) 

# and confidence? yeah sure.
count(viirs_notwf_f, CONFIDENCE, anyPolygons)
viirs_notwf_f <- filter(viirs_notwf_f, CONFIDENCE != 'l')

mapview_points <-viirs_notwf_f %>% 
  st_as_sf(sf_column_name = 'geometry') %>% 
  mapview(zcol = "jday_adj", cex = 2, alpha = "anyPolygons") 

# fires
pal <- c("#d1495b", "#edae49", "#00798c")
mapview_fires <- 
  mapview(fires$frap_rx, col.regions = pal[1], color = pal[1], legend = F) +
  mapview(fires$CalMAPPER, col.regions = pal[2], color = pal[2], legend = F) +
  mapview(fires$facts, col.regions = pal[3], color = pal[3], legend = F) 

# land ownership, ag
SRA_raster <- raster(rasts$SRA)
nonag_raster <- raster(rasts$non_ag)

# map it all!
mapview(SRA_raster, alpha = .2) +
  mapview(nonag_raster, alpha = .2, legend = F)+
  mapview_fires + 
  mapview_points

```

How many of these filtered points are within the polygons? And by ownership?
```{r}
# points that intersect treatments vs not
viirs_notwf_f %>% 
  filter(!is.na(SRA)) %>% 
  group_by(anyPolygons) %>% 
  summarise(n = n()) %>% 
  rename(intersects_trts = "anyPolygons")

# points that intersect treatments vs not AND responsibility areas (state, local, federal)
viirs_notwf_f %>% 
  filter(!is.na(SRA)) %>% 
  group_by(anyPolygons, SRA) %>% 
  summarise(n = n()) %>% 
  rename(intersects_trts = "anyPolygons")
```

