---
title: "Q2_VIIRS_wo_records"
author: "Lisa Rosenthal"
date: "2023-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(fs)
library(terra)
library(mapview)
library(fasterize)

```

### Functions
```{r}

# manipulations dont like it when geometries aren't valid. and they don't work with NA validity.
make_valid <- function(sf_object) {
  validity <- st_is_valid(sf_object)
  if (sum(is.na(validity)) > 0) {
    stop("can't proceed. Contains rows where validity is NA.")
  } else{
    if (all(validity)) {
      return(sf_object)
    } else{
      sf_object <- st_make_valid(sf_object)
      print(glue::glue("there were {sum(validity == F)} not valid geometries"))
      return(sf_object)
    }
  }
}

# transform to VIIRS crs, select fields you want, make valid
prep_polys <- function(polygon, columns, points = viirs){
  message("transforming polygons")
  polygon <- polygon %>% 
  st_transform(st_crs(points)) %>% 
  select(all_of(columns))
  
  message("making polygons valid")
  make_valid(polygon)
}


# extract data. polygons shouldn't overlap, or it wont run.
extract_polys <- function(polygons, points = viirs) {
  intersects <- st_intersects(points, polygons, sparse = T)
  if ( length(unlist(intersects)) == 0)
    return(warning("No intersections"))
  else{
    intersections_list <- map(intersects, unlist)
    
    # error if there's more than 1 polygon intersecting a point
    if(length(unlist(intersections_list)) > nrow(points)){
     stop("Point(s) overlapped with more than 1 polygon. Not going to run.")
    # make sure points with no intersection become NA
      }else{
      polygonIDs <- map_vec(intersections_list, 
                           ~ ifelse(sum(.x) == 0, NA, .x) )
      newcols <- st_drop_geometry(polygons)[polygonIDs, ]
      return(newcols)
    }
  }
}

```

## Goal

Create a cleaned up version of VIIRS points consisting of high confidence fire points. Was going to turn the points into circles, since it does have a resolution of 375m, but for this level of analysis, I'm keeping them as points. Computation is going to be a lot faster and I doubt I'm introducing that much bias. 

Remove points that are
- On ag land (already did that with cropland data layer, but use a CA specific one)
- campgrounds
- power plants
- solar farms
- water
- intersect at all with urban land. can create stationary hotspots.

Other layers that could be important
- Jurisdictions/Land management agency

For polygon layers that are wall-to-wall, convert them to rasters. For rasters in general, will want to resample at a resolution of 375 m. For categorical values, use the mode (most common pixel) when aggregating. 

to rid stationary hotspots, could rasterize the points and calculate the density of viirs points during the non-rx fire months. >4 unique days get thrown out? 

## Load CA boundary and VIIRS data

```{r}
# California boundary
ca_path <- file.path(ref_path, 'CA boundary/ca-state-boundary/CA_State_TIGER2016.shp')
CA <- st_read(ca_path) 

# VIIRS fire detects
viirs_dir <- file.path(ref_path, 'VIIRS/CA_2020')
viirs_shps <- dir_ls(viirs_dir, recurse = T, glob = '*.shp')
viirs <- map_df(viirs_shps, st_read) %>% st_transform(st_crs(CA))
viirs <- viirs[CA,]

# vis extent of the two. reminder that I need to redownload VIIRS to completely cover the southern part of CA
st_make_grid(st_bbox(CA), n = 1) %>% plot(col = 'maroon4')
st_make_grid(st_bbox(viirs), n = 1) %>% plot(add = T, col = 'tan')
plot(st_geometry(CA), add = T)


```

## Vector based layers

Layers you might want are...
- counties, 
- responsibility areas, 
- campgrounds
- power plants
- solar farms
- land/water

Will want to convert the wall-to-wall layers to rasters. Will be easier to deal with in the future. Use a raster with VIIRS' resolution--375m. 

```{r}
# wall to wall data

# responsibility areas--local, state, federal
RA_path <- file.path(ref_path, 'Responsibility areas') %>% 
  dir_ls(recurse = T, glob = "*shp")
RA <- st_read(RA_path)
# make field of interest integer
#  SRA   SRA_int
# <chr>   <int>
# 1 FRA         1
# 2 LRA         2
# 3 SRA         3
RA <- RA %>% mutate(SRA_int = as.integer(as.factor(SRA)), 
              .after = SRA) 

# USFS
NF_CA <- st_read(file.path(ref_path, "National Forest Boundaries/USFS_CA.shp"))
NF_CA

# template raster to rasterize to
raster_pars <- list(
  crs = st_crs(viirs)$epsg, 
  extent = raster::extent(st_bbox(viirs)),
  res = c(375, 375)
)
raster_temp <- with(raster_pars, 
                    raster::raster(ext = extent, resolution = res, crs = crs))

getmode <- function(v,na.rm) {
uniqv <- unique(v)
uniqv[which.max(tabulate(match(v, uniqv)))]
} 
getmode(c(1,2,2,2,2,1))

RA_r <- fasterize(RA, raster_temp, field = 'SRA_int', by = "SRA")
RA_r <- rast(RA_r)

plot(RA_r)
hist(RA_r$LRA)
plot(RA_r$LRA)
values(RA_r$LRA, na.rm = T)
spatSample(RA_r$LRA, 4, na.rm = T)

RA_rast(RA_r)

plot(RA_r)
hist(RA_r)
```



```{r}
# load data ==============================================

# counties
counties <- file.path(ref_path, "Counties") %>% 
  dir_ls(recurse = T, glob = "*.shp") %>% 
  st_read



# campgrounds
camping_FS <- file.path(ref_path, 'camping/USFS') %>% 
  dir_ls(recurse = T, glob = '*shp') %>% 
  st_read %>% 
  st_transform(st_crs(viirs))
camping_CA <- file.path(ref_path, 'camping/CA_state_parks') %>% 
  dir_ls(recurse = T, glob = '*shp') %>% 
  st_read %>% 
  st_transform(st_crs(viirs))
camping_FS <- filter(camping_FS, grepl('camp', PARENTACTI, ignore.case = T))
camping_FSCA <- camping_FS[CA,]
camping <- bind_rows(
  transmute(camping_FSCA, camping = 'USFS'),
  transmute(camping_CA, camping = "CA")
)

# power plants--not sure how to filter these. any power plants have gas flares, even solar.
# larry hunsinger
powerplants <- file.path(ref_path, 'powerplants') %>% dir_ls(glob = "*geojson") %>% st_read
powerplants$PriEnergySource %>% unique %>% sort %>% cbind
powerplants$PlantName %>% unique %>% sort


# functions to extract data at VIIRS points ============

# transform to VIIRS crs, select fields you want, make valid
counties_valid <- prep_polys(counties, c("NAME", 'NAMELSAD'))
RA_valid <- prep_polys(RA, "SRA")
camping_valid <- prep_polys(camping, "camping") %>% st_buffer(dist = 375/2)


# extract data. polygons shouldn't overlap, or it wont run.
counties_ex <- extract_polys(counties_valid, viirs) 
RA_ex <- extract_polys(RA_valid, viirs) 
camping_ex <- extract_polys(camping_valid, viirs)





```
## Raster layers

```{r}
# load data ===============
# NASS crop layer
cdl_path <- file.path(ref_path, "NASS/CDL/2020_30m_cdls") %>% dir_ls(glob = '*.tif')
cdl_NASS <- rast(cdl_path)

# crop to CA
f_crop_to_CA <- function(raster){
  CA_t <- st_transform(CA, crs(raster))
  raster_CA <- crop(raster, CA_t)
  return(raster_CA)
}
cdl_NASS_CA <- f_crop_to_CA(cdl_NASS)

# create a mask by identifying classes to keep
levels_table <- levels(cdl_NASS_CA)[[1]]
# levels_table$Class_Names
nonAgVeg <- c('Pasture/Grass', 'Forest', "Shrubland", "Barren", "Christmas Trees",
              "Nonag/Undefined", "Deciduous Forest", "Evergreen Forest", "Mixed Forest",
              "Grassland/Pasture", "Woody Wetlands", "Herbaceous Wetlands"
              )
nonAgVeg_values <- levels_table$value[levels_table$Class_Names %in% nonAgVeg]

# reclassify the raster to be non-ag vs. not
mat1 <- cbind(nonAgVeg_values, 1)
cdl_CA_mask <- classify(cdl_NASS_CA, mat1, others = 0)
plot(cdl_CA_mask)

# extract with VIIRS
viirs_t <- st_transform(viirs, crs(cdl_NASS_CA))
vals <- extract(cdl_CA_mask, viirs_t)
viirs$cdl_notAg <- vals$Class_Names

```



```{r}

# TRASH

counties_t <- st_transform(counties, st_crs(RA))
county_filter <- filter(counties_t, NAME == 'Sacramento')
RA_filter <- RA[county_filter,]
st_intersects(RA, county_filter)
st_intersection(RA, county_filter)

# png(file.path(ref_path, "Responsibility areas/State_Responsibility_Areas.gdb", 'SRA.png'),
#     width = 6, height = 4, units = 'in', res = 72)
# plot(RA['SRA'])
# dev.off()

# # CNRA crop layer -- way too complicated
# crop_path <- file.path(ref_path, "CNRA/crop_mapping") %>% 
#   dir_ls(glob = "*2020*.gdb", recurse = T)
# crop_CNRA <- st_read(crop_path, layer = 'i15_Crop_Mapping_2020') 
# head(crop_CNRA)
# cols <- c("SYMB_CLASS", "CLASS1", "CLASS2", "CLASS3", "MAIN_CROP")
# map(cols, ~ unique(crop_CNRA[[.x]])) %>% set_names(cols)

```



```{r}
viirs_dir <- file.path(ref_path, 'VIIRS/CA_2020')
viirs_nonAg <- read_rds( file = file.path(viirs_dir, 'viirs_nonAg.rds'))
viirs_nonAg

viirs_nonAg$SCAN %>% hist
```

