---
title: "Q3_projects_vs_time"
author: "Lisa Rosenthal"
date: "2023-09-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# devtools::install_github("r-spatial/mapview@develop")
library(mapview)
library(tidyverse)
library(sf)
library(fs)
library(units)

theme_set(theme_bw() + theme(panel.grid.minor = element_blank()))
```

# Question 3

For burn projects that took more than a week to wrap up: How well did the number of days for which thermal anomalies were detected corroborate with the number of days that burners indicated in their project reporting? How does the level of corroboration change as a burn project stretches out longer in time?

## Approach

For every day an activity occurs, assess if there was a VIIRS detection within its treatment polygon. Compare % of temporally correct detections to duration of activity. 

## Prep data

Load VIIRS and CalMAPPER data. CalMapper polygons buffered with a radius of half the VIIRS resolution.

```{r}
# VIIRS fire detects
viirs <- read_rds(file.path(ref_path, 'VIIRS/CA_2020/viirs_nonAg.rds')) 


# load up CalMAPPER data, created in Q1
CM_burn2020_buffered <- read_rds(file.path(ref_path, 'CalMAPPER/burnactivities_2020_buffered.rds'))
CM_trtIDs_buffered <- read_rds(file.path(ref_path, 'CalMAPPER/treatmentgeometries_2020_buffered.rds'))
```

Get VIIRS points that intersect with the polygons.

```{r}
# get the viirs points contained in the treatment polygons
viirs_t <- st_transform(viirs, st_crs(CM_burn2020_buffered))

# find those that intersect, do with each activity
viirs_int <- viirs_t %>% 
  filter(CONFIDENCE != 'l') %>% 
  st_intersection(CM_burn2020_buffered) %>% 
  mutate(rowname = rownames(.), 
         pointID = floor(as.numeric(rowname)),
         overlap_trts = grepl('\\.', rowname),
         .before = LATITUDE) %>% 
  group_by(pointID) %>% 
  mutate(overlap_trts = any(overlap_trts == T)) %>% 
  ungroup() 
#print(viirs_int %>% arrange(AQ_ID), width = Inf)

  
```


For every activity, calculate if VIIRS points were detected during CalMAPPER activity. how many points were during the recorded activity? how many detected but outside the activity window?

```{r}
# each row is a viirs point that intersected a polygon. A single VIIRS point can appear >1 if it intersects more than one activity. Likewise, an activity can appear >1 if there are multiple VIIRS points intersecting it. 
viirs_matches <- viirs_int %>% 
  mutate(date_matches_CM = date >= ACTIVITY_START & date <= ACTIVITY_END, 
         dateadj_matches_CM = date_adj >= ACTIVITY_START & date_adj <= ACTIVITY_END, 
         dateadj_matches_CM_buffer = (date_adj + 1) >= ACTIVITY_START &
           (date_adj - 1) <= ACTIVITY_END,
         .after = date) 


# group by activity and calculate VIIRS detections overlapping it
mean(yday(viirs_matches$ACTIVITY_START[1]), yday(viirs_matches$ACTIVITY_END[1]))
matches_activitylevel <- viirs_matches %>% 
  as_tibble() %>% 
  mutate(CM_jday = (yday(ACTIVITY_START) + yday(ACTIVITY_END)) / 2) %>% 
  group_by(AQ_ID, TREATMENT_ID, CM_jday) %>% 
  summarise(duration = as.numeric(duration[1]),
            trt_acres = as.numeric(trt_acres[1]),
            unique_viirs_days = length(unique(date)),
            n_matching_dates = length(unique(date[date_matches_CM == T])),
            n_matching_dates_adj = length(unique(date_adj[dateadj_matches_CM == T])), 
            n_matching_dates_adj_buffer = length(unique(date_adj[dateadj_matches_CM_buffer == T])), 
            viirs_jday = mean(unique(jday))
            ) %>% 
  arrange(TREATMENT_ID) %>% 
  ungroup

# check it out
print(matches_activitylevel, width = Inf)
matches_activitylevel %>% 
  select(n_matching_dates:n_matching_dates_adj_buffer) %>% 
  apply(., 2, function(x) sum(x > 0))

# how many activities match with the buffer, but not exact viirs day? 
# 173 activites don't match either exact or buffered time (212 if exact). 86 match exactly, 39 additional ones with the buffer.
viirs_matches %>% 
  as_tibble %>% 
  group_by(AQ_ID) %>% 
  summarise(matches_exact = sum(dateadj_matches_CM) > 0,
            matches_buffer = sum(dateadj_matches_CM_buffer) > 0) %>% 
  select(-AQ_ID) %>% table
(39 + 86) / 86 # how many more when you buffer the date? 
125 / length(unique(CM_burn2020_buffered$AQ_ID)) # brings it up to 18% instead of 12%

# look at duration according to CM vs viirs
matches_activitylevel %>% 
  mutate(diff = as.numeric(duration) - unique_viirs_days) %>% 
  ggplot(aes(diff)) +
  geom_rect(aes(xmin = -Inf, xmax = 0, ymin = -Inf, ymax = Inf), fill = "lightcyan2") +
  geom_rect(aes(xmin = 0, xmax = Inf, ymin = -Inf, ymax = Inf), fill = "indianred1") +
  geom_histogram(color = 'grey20', fill = 'grey70', alpha = .9) +
  geom_vline(xintercept = 0, lty = 2) +
  labs(x = "Difference in # of days per activity: \nCalMAPPER - VIIRS", 
       y = "# of activities") +
  scale_x_continuous(n.breaks = 10)

# look at average julian day according to CM vs VIIRS
matches_activitylevel %>% 
  mutate(diff = CM_jday  - viirs_jday) %>% 
  ggplot(aes(diff)) +
  geom_rect(aes(xmin = -Inf, xmax = 0, ymin = -Inf, ymax = Inf), fill = "lightcyan2") +
  geom_rect(aes(xmin = 0, xmax = Inf, ymin = -Inf, ymax = Inf), fill = "indianred1") +
  geom_histogram(color = 'grey20', fill = 'grey70', alpha = .9) +
  geom_vline(xintercept = 0, lty = 2) +
  labs(x = "Difference in avg Julian day per activity: \nCalMAPPER - VIIRS", 
       y = "# of activities") +
  scale_x_continuous(n.breaks = 10)


# look at % of activity days with VIIRS detections vs. duration
ggplot(matches_activitylevel, 
       aes(as.numeric(duration), 100 * n_matching_dates/as.numeric(duration))) + 
  geom_point() + 
  labs(x = 'Activity duration (days)', 
       y = "% activity days with VIIRS detections") +
  scale_y_continuous(limits = c(0, 100))

# same thing as above but x-axis is activity size
matches_activitylevel %>%
  filter(n_matching_dates > 0 ) %>%
  ggplot(aes(trt_acres, 100 * n_matching_dates/as.numeric(duration))) +
  geom_point() +
  labs(x = 'Polygon size (acres)',
       y = "% activity days with VIIRS detections") +
  scale_y_continuous(limits = c(0, 100)) #+ facet_grid(~ duration <= 7)

# # probability of any VIIRS detections vs. duration
# ggplot(matches_activitylevel, aes(duration, as.numeric(n_matching_dates != 0))) +
#   geom_point(alpha = .2) +
#   geom_smooth(method="glm", method.args = list(family = "binomial")) +
#   labs(x = 'Activity duration (days)', 
#        y = "Probability of any VIIRS detections")

# % of activities with no VIIRS detections on those days
mean(matches_activitylevel$n_matching_dates == 0)



# number of days spatially overlapping that do not temporally match
# i.e. the days with satellite detections without calMAPPER records 
# --> false positives or missing calMAPPER records
matches_activitylevel %>% 
  mutate(any_detection = as.numeric(n_matching_dates != 0),
         dur_class = cut(duration, breaks = seq(0, max(duration) + 7, by = 7), 
                         labels = F)) %>% 
  group_by(dur_class) %>% 
  summarise(Yes = sum(any_detection), 
            No = n() - Yes) %>%
  pivot_longer(-dur_class) %>% 
  ggplot(aes(dur_class, value, group = name, fill = name)) +
  geom_col() +
  labs(x = "Activity duration (weeks)", 
       y = "# of activities", 
       fill = "VIIRS detected \nduring activity", 
       title = "When there's spatial overlap, how many activities temporally overlap
       with at least 1 VIIRS detection?") +
  scale_x_continuous(n.breaks = 10)



```


Those were just the activities where VIIRS intersecting. how many activities without VIIRS detections? if there were detections (spatial overlap with trt), how many didn't temporally match? Make a pie chart. 

```{r}

spatial <- c('No spatial overlap', 
           "Spatial overlap", "Spatial overlap")
temporal <- c("No temporal match", "No temporal match", "Temporal match")
values <- c(
  length(unique(CM_burn2020_buffered$AQ_ID)) - length(unique(matches_activitylevel$AQ_ID)),
  sum(matches_activitylevel$n_matching_dates == 0),
  sum(matches_activitylevel$n_matching_dates > 0)
)
dat <- tibble(spatial, temporal, values) %>% 
  mutate(percent = round(values/sum(values)*100, 2))
print(dat)

# viz that breakdown
treemap::treemap(dat, index = c("spatial", 'temporal'), vSize = "values", type = "index",
                 align.labels=list(
        c("center", "center"), 
        c("center", "bottom")
        ),                                   
    overlap.labels = .5, 
    inflate.labels = F,
    border.col=c("black","white"),
    border.lwds=c(3,2),
    fontsize.labels=c(15,12),
    palette = 'Set1',
    title = "CalMAPPER activities: Spatial + temporal overlap with VIIRS")



```


Supply Anny with a few tables---activities that spatially overlap with VIIRS and information on how many of the detects actaully temporally occur during the recorded window.

```{r}

# for every activity, return activity data and intersecting viirs dates
for_anny <- viirs_matches %>% 
  as_tibble() %>% 
  select(viirs_date = date_adj, AQ_ID, ACTIVITY_START, ACTIVITY_END, 
         matches = dateadj_matches_CM) %>% 
  mutate(ACTIVITY_START = date(ACTIVITY_START),
         ACTIVITY_END = date(ACTIVITY_END)) %>% 
  arrange(AQ_ID, viirs_date)

for_anny2 <- viirs_matches %>% 
  distinct(AQ_ID, date, .keep_all = T) %>% 
  as_tibble() %>% 
  group_by(AQ_ID) %>% 
  summarise(n_viirs_days = length(unique(date)), 
            n_matches = sum(dateadj_matches_CM), 
            activ_dur = duration[1])

write_csv(for_anny, 'tmp_outputs/CMactivity_viirs_days.csv')
write_csv(for_anny2, 'tmp_outputs/CMactivity_viirs_days_summary.csv')

```

Another thing I want to do is to check if there are any VIIRS points that are spatially close to CalMAPPER polygons, but not inside. could the polygons be mapped incorrectly? 

```{r}

# make the calmapper polygon data easier to view on mapviewer. 
unique_values <- function(values){
  vals <- values %>% unique %>% sort
  paste(vals, collapse = ", ")
}
CM_trt_simplified <- CM_burn2020_buffered %>% 
  group_by(TREATMENT_ID) %>% 
  summarise(AGENCY_NAME = unique(AGENCY_NAME), 
            n_broadcast = sum(broadcast_burn), 
            n_pile = sum(pile_burn),
            start_months = unique_values(month(ACTIVITY_START)),
            end_months = unique_values(month(ACTIVITY_END)),
            acres_recorded = sum(TREATED_ACRES),
            acres_polygon = unique(trt_acres), 
            veg = unique_values(BROAD_VEG_TYPE)
            )

# identify polygons with activities that spatially and temporally match (buffered = +/-1 day)
all_trtIDs <- unique(CM_burn2020_buffered$TREATMENT_ID)
temporal_matches <- matches_activitylevel %>% 
  group_by(TREATMENT_ID) %>% 
  summarise(buffered_temp_match = sum(n_matching_dates_adj_buffer > 0) / n()) %>% 
  mutate(spatial_match = 1)
trt_overlaps <- bind_rows(
  temporal_matches, 
  tibble(TREATMENT_ID = all_trtIDs[!all_trtIDs %in% temporal_matches$TREATMENT_ID], 
       buffered_temp_match = 0,
       spatial_match = 0)
)
CM_trt_simplified <- left_join(CM_trt_simplified, trt_overlaps)




# constrain to a smaller area. choose some county that has fire. 
counties <- file.path(ref_path, "Counties") %>% 
  dir_ls(recurse = T, glob = "*.shp") %>% 
  st_read %>% 
  st_transform(st_crs(CM_burn2020_buffered))

sort(counties$NAME)
mapview(counties, zcol = "NAME")
counties_filter <- filter(counties, 
                          NAME %in% c('Nevada', "Sierra", "Placer", "El Dorado"))

# filter viirs and CalMAPPER
viirs_filter <- viirs_t[counties_filter,] #%>% filter(SATELLITE == 'N') # just suomi
CM_filtered <- CM_trt_simplified[counties_filter,]
mapview(CM_filtered,
        label= "TREATMENT_ID",
        lwd = 1,
        zcol =  "buffered_temp_match",
        legend = F
        ) +
mapview(CM_filtered,
        label= "TREATMENT_ID",
        lwd = 1,
        zcol = "spatial_match",
        legend = F
        ) +
  mapview(viirs_filter, zcol = "jday_adj", 
          at = seq(0, 365, 30),
          cex = 3, alpha = .5)
  

# look at particular treatments and their overlap
viirs_int %>% 
  filter(TREATMENT_ID == 10572) %>% 
  print(width = Inf)

```


