---
title: "TTS_acreage"
output: html_document
date: "2023-11-16"
---

```{r, include = F}
require(tidyverse)
require(sf)
require(fs)
require(mapview)
require(writexl)
```

# Load data
```{r}
gdb <- "C:/Users/ctubbesi/OneDrive - California Air Resources Board/Documents/Reference data/Treatment Tracking/ITS Fire 20230926.gdb"
```

```{r}
st_layers(gdb)
pts <- st_read(gdb, layer = 'ITS_Fire_Points_20230822')
pols <- st_read(gdb, layer = 'ITS_Fire_Polygons_20230822')
```

# Save as tables for Anny
```{r}

# Assuming you have two data frames: df1 and df2
# Save df1 as "Sheet1" and df2 as "Sheet2" in the same Excel file
write_xlsx(list(Sheet1 = df1, Sheet2 = df2), "myfile.xlsx")

write_xlsx(list(Points = pts %>% st_drop_geometry(), Polygons = pols %>% st_drop_geometry()), "excel/ITS_Fire_20230822.xlsx")
```

# Filter to records that are complete and that are measured in acres
```{r}
pts <- pts %>% 
  filter(ACTIVITY_STATUS == "COMPLETE") %>% 
  filter(ACTIVITY_UOM == "AC")
pols <- pols %>% 
  filter(ACTIVITY_STATUS == "COMPLETE")
```

# Remove wildfire for resource benefit
```{r}
summary(as.factor(pts$ACTIVITY_DESCRIPTION))
summary(as.factor(pols$ACTIVITY_DESCRIPTION))
```

```{r}
pols <- pols %>% 
  filter(!ACTIVITY_DESCRIPTION %in% c("WM_RESRC_BENEFIT", "PL_TREAT_BURNED"))
```

# Are there Cal Fire pile burns?
```{r}
pols %>% 
  st_drop_geometry() %>% 
  group_by(AGENCY, ADMINISTERING_ORG, IMPLEMENTING_ORG, ACTIVITY_DESCRIPTION) %>% 
  filter(IMPLEMENTING_ORG == "CALFIRE" & ACTIVITY_DESCRIPTION == "PILE_BURN") %>% 
  View()
```


# How many PFIRS records in pts?
```{r}
pts %>% group_by(Source, Year) %>% 
  count()
```

# Sum acreage
```{r}
pts_acreage <- pts %>% 
  st_drop_geometry() %>% 
  group_by(Year) %>% 
  summarize(ac_pts = sum(na.omit(ACTIVITY_QUANTITY)))
pts_acreage
```

```{r}
pols_acreage <- pols %>% 
  st_drop_geometry() %>% 
  group_by(Year) %>% 
  summarize(ac_pols = sum(na.omit(ACTIVITY_QUANTITY)))
pols_acreage
```



```{r}
tts_ac <- full_join(pts_acreage, pols_acreage)
tts_ac <- tts_ac %>% 
  mutate(tts_acres = ac_pts + ac_pols)
tts_ac
```

# Save table as Csv
```{r}
write.csv(tts_ac, file = "excel/tts_acres_yr.csv", row.names = F)
```

# Look for USFS data
```{r}
pts %>% st_drop_geometry() %>% group_by(Source, PRIMARY_OWNERSHIP_GROUP, ADMINISTERING_ORG, IMPLEMENTING_ORG, IMPLEM_ORG_NAME, ADMIN_ORG_NAME) %>% count() %>% View()
```

# Look for private industrial timberland data
```{r}
pts %>% 
  st_drop_geometry() %>% 
  filter(IMPLEMENTING_ORG == "Sierra Pacific Industries") %>% 
  group_by(PRIMARY_OWNERSHIP_GROUP, ACTIVITY_DESCRIPTION, Year) %>% 
  summarize(acres = sum(ACTIVITY_QUANTITY))
```

# Check for records NOT on private industrial timberland bit that were clearly private industrial
```{r}
pts %>% 
  st_drop_geometry() %>% 
  filter(PRIMARY_OWNERSHIP_GROUP != "PRIVATE_INDUSTRY") %>% 
  select(IMPLEMENTING_ORG, IMPLEM_ORG_NAME) %>% 
  distinct() %>% 
  as.list()
```

# Create a new column for burns on private industrial land or conducted by SPI
```{r}
pts_industrial <- pts %>% 
  st_drop_geometry() %>% 
  filter(PRIMARY_OWNERSHIP_GROUP == "PRIVATE_INDUSTRY" | IMPLEMENTING_ORG %in% c("Sierra Pacific Industries", "Michigan-California Timber Company")) %>% 
  mutate(type = "points")

pts_industrial %>% 
  group_by(PRIMARY_OWNERSHIP_GROUP, ACTIVITY_DESCRIPTION, Year, Source) %>% 
  summarize(ac = sum(ACTIVITY_QUANTITY)) %>% 
  print(n = 33)
```

```{r}
pols_industrial <- pols %>% 
  st_drop_geometry() %>% 
  filter(PRIMARY_OWNERSHIP_GROUP == "PRIVATE_INDUSTRY" | IMPLEMENTING_ORG %in% c("Sierra Pacific Industries", "Michigan-California Timber Company")) %>% 
  mutate(type = "polygon")

pols_industrial %>% 
  group_by(PRIMARY_OWNERSHIP_GROUP, IMPLEMENTING_ORG, ACTIVITY_DESCRIPTION, Year, Source) %>% 
  summarize(ac = sum(ACTIVITY_QUANTITY)) %>% 
  print(n = 33)
```

## Combine industrial polygons and industrial points
```{r}
industrial <- full_join(pts_industrial, pols_industrial) %>% 
    group_by(PRIMARY_OWNERSHIP_GROUP, IMPLEMENTING_ORG, ACTIVITY_DESCRIPTION, Year, Source) %>% 
  summarize(count = n(), acres = sum(ACTIVITY_QUANTITY)) %>% 
  select(Source, everything()) %>% 
  filter(Year != 2023)
industrial
```

```{r}
industrial_summary <- industrial %>% 
  group_by(Source, Year) %>% 
  summarize(count = n(), acres = sum(acres))
industrial_summary
```


# Save them
```{r}
write_xlsx(list(full = industrial, summary = industrial_summary), "excel/industrial_timber_acreage.xlsx")

```

# DOE
```{r}
pts %>% 
  st_drop_geometry() %>% 
  filter(IMPLEMENTING_ORG == "US Department of Energy") %>% 
  group_by(PRIMARY_OWNERSHIP_GROUP) %>% 
  select(IMPLEMENTING_ORG, PRIMARY_OWNERSHIP_GROUP, ACTIVITY_QUANTITY, ACTIVITY_UOM, Year) %>% 
  group_by(Year) %>% 
  summarize(sum(na.omit(ACTIVITY_QUANTITY)))
```

