---
title: "TTS_date_ranges"
output: html_document
date: "2023-11-28"
---


```{r, include = F}
require(tidyverse)
require(sf)
require(fs)
require(mapview)
```

# Load data
```{r}
gdb <- "C:/Users/ctubbesi/OneDrive - California Air Resources Board/Documents/Reference data/Treatment Tracking/ITS Fire 20230926.gdb"
gdb_public <- "C:/Users/ctubbesi/OneDrive - California Air Resources Board/Documents/Reference data/Treatment Tracking/ITS_Dashboard_Nov2023.gdb"
```

```{r}
st_layers(gdb)
pts <- st_read(gdb, layer = 'ITS_Fire_Points_20230822')
pols <- st_read(gdb, layer = 'ITS_Fire_Polygons_20230822')
```

```{r}
st_layers(gdb_public)
pts_public <- st_read(gdb_public, layer = 'ITS_Points_20230822')
pols_public <- st_read(gdb_public, layer = 'ITS_Polygons_20230822')
pts_public <- pts_public %>% 
  filter(activity_cat == "BENEFICIAL_FIRE")
pols_public <- pols_public %>% 
  filter(activity_cat == "BENEFICIAL_FIRE")

```


# Filter to records that are complete and that are measured in acres
```{r}
pts <- pts %>% 
  filter(ACTIVITY_STATUS == "COMPLETE") %>% 
  filter(ACTIVITY_UOM == "AC")
pols <- pols %>% 
  filter(ACTIVITY_STATUS == "COMPLETE")
```

# Remove wildfire for resource benefit
```{r}
summary(as.factor(pts$ACTIVITY_DESCRIPTION))
summary(as.factor(pols$ACTIVITY_DESCRIPTION))
```

```{r}
pols <- pols %>% 
  filter(!ACTIVITY_DESCRIPTION %in% c("WM_RESRC_BENEFIT", "PL_TREAT_BURNED"))
```


# Filter out 2023 since it's preliminary
```{r}
pts <- pts %>% 
  filter(Year != 2023)
pols <- pols %>% 
  filter(Year != 2023)
```

# Add DURATION columns
```{r}
pts <- pts %>% 
  mutate(ACTIVITY_START = as.Date(ACTIVITY_START)) %>% 
  mutate(ACTIVITY_END = as.Date(ACTIVITY_END)) %>% 
  mutate(DURATION = ACTIVITY_END - ACTIVITY_START+1)
```

```{r}
pols <- pols %>% 
  mutate(ACTIVITY_START = as.Date(ACTIVITY_START)) %>% 
  mutate(ACTIVITY_END = as.Date(ACTIVITY_END)) %>% 
  mutate(DURATION = ACTIVITY_END - ACTIVITY_START+1)
```

## Assign duration=1 if data source is PFIRS

```{r}
pts %>% 
  group_by(Source, DURATION) %>% 
  count()
```

```{r}
pts$DURATION[is.na(pts$DURATION)] <- ddays(1)/86400
```

```{r}
pts %>% 
  group_by(Source, DURATION) %>% 
  count()
```

```{r}
pts %>% 
  select(PROJECTID_USER, AGENCY, ADMINISTERING_ORG, COUNTY, ACTIVID_USER, TRMTID_USER, ACTIVITY_NAME, ACTIVITY_START, ACTIVITY_END, ACTIVITY_QUANTITY, DURATION) %>% 
  filter(AGENCY == "DOI") %>% 
  arrange(desc(ACTIVITY_QUANTITY))
```

# See if public data has the same issues
```{r}
pols_public %>% 
  filter(agency == "DOI") %>% 
  arrange(desc(activity_quantity))
```

# Invert duration for DOI records
```{r}
pols <- pols %>% 
  mutate(DURATION = case_when(
    AGENCY == "DOI" ~ DURATION*(-1),
    TRUE ~ DURATION
  ))
```

# Sum acreage
```{r}
pts_acreage <- pts %>% 
  st_drop_geometry() %>% 
  group_by(DURATION) %>% 
  summarize(ac_pts = sum(na.omit(ACTIVITY_QUANTITY)))
pts_acreage
```

```{r}
pols_acreage <- pols %>% 
  st_drop_geometry() %>% 
  group_by(DURATION) %>% 
  summarize(ac_pols = sum(na.omit(ACTIVITY_QUANTITY))) 
pols_acreage
```

```{r}
ac_both <- full_join(pts_acreage, pols_acreage) 
ac_both$ac_pts[is.na(ac_both$ac_pts)] <- 0
ac_both$ac_pols[is.na(ac_both$ac_pols)] <- 0
ac_both <- ac_both %>% 
  mutate(ac_both = ac_pts + ac_pols) 
ac_both
```

## Bin it
```{r}
acres_ITS <- ac_both %>%
  mutate(duration_group = case_when(
    DURATION == 1 ~ "1 day",
    DURATION == 2 ~ "2 days",
    DURATION >= 3 & DURATION <= 7 ~ "3-7 days",
    DURATION >= 8 & DURATION <= 31 ~ "8-31 days",
    DURATION > 31 & DURATION <= 180 ~ "32-180 days",
    DURATION > 180 ~ ">180 days"
  )) %>%
  group_by(duration_group) %>%
  summarise(ac_both = sum(ac_both, na.rm = TRUE))
acres_ITS <- acres_ITS %>% 
  mutate(percent = ac_both/sum(ac_both)*100) %>% 
  t()
acres_ITS
```

## Write table
```{r}
write.csv(acres_ITS, file = "excel/ITS_duration_ac.csv")
```

# Count records
```{r}
pts_count <- pts %>% 
  st_drop_geometry() %>% 
  group_by(DURATION) %>% 
  count() %>% 
  rename(n_pts = n)
pts_count
```

```{r}
pols_count <- pols %>% 
  st_drop_geometry() %>% 
  group_by(DURATION) %>% 
  count() %>% 
  rename(n_pols = n)
pols_count
```

```{r}
n_both <- full_join(pts_count, pols_count) 
n_both$n_pts[is.na(n_both$n_pts)] <- 0
n_both$n_pols[is.na(n_both$n_pols)] <- 0
n_both <- n_both %>% 
  mutate(n_both = n_pts + n_pols) 
n_both
```

## Bin it
```{r}
duration_ITS <- n_both %>%
  mutate(duration_group = case_when(
    DURATION == 1 ~ "1 day",
    DURATION == 2 ~ "2 days",
    DURATION >= 3 & DURATION <= 7 ~ "3-7 days",
    DURATION >= 8 & DURATION <= 31 ~ "8-31 days",
    DURATION > 31 & DURATION <= 180 ~ "32-180 days",
    DURATION > 180 ~ ">180 days"
  )) %>%
  group_by(duration_group) %>%
  summarise(n_both = sum(n_both, na.rm = TRUE))
duration_ITS <- duration_ITS %>% 
  mutate(percent = n_both/sum(n_both)*100) %>% 
  t()
duration_ITS
```

## Write table
```{r}
write.csv(duration_ITS, file = "excel/ITS_duration_count.csv")
```

```{r}
pts_ac_plot <- pts_acreage %>% 
  rename(ac = ac_pts) %>% 
  mutate(source = "pts")
pols_ac_plot <- pols_acreage %>% 
  rename(ac = ac_pols) %>% 
  mutate(source = "pols")
both_ac_plot <- full_join(pts_ac_plot, pols_ac_plot) %>% 
  filter(!is.na(DURATION))
both_ac_plot
```




```{r}
ggplot(both_ac_plot)+
  geom_col(aes(y = ac, x = DURATION, fill = source))+
  theme_minimal()+
  xlab("Duration (days)")
  scale_x_continuous(breaks = c(0, 2.3, 4.6, 6.9), labels = c("1", "10", "100", "1000"))
```


# Histogram

## Combine raw tables
```{r}
pts_join <- pts %>% 
  st_drop_geometry() %>% 
  mutate(source = "Points") 
pols_join <- pols %>% 
  st_drop_geometry() %>% 
  mutate(source = "Polygons")
both <- full_join(pts_join, pols_join)
nrow(both)==nrow(pts) + nrow(pols)
```

```{r}
ggplot(both)+
  geom_histogram(aes(x = log(as.numeric(DURATION)), fill = source), position = "dodge", binwidth = 0.3)+
  theme_minimal()+
  xlab("Duration (days)")+
  scale_x_continuous(breaks = c(0, 2.3, 4.6, 6.9), labels = c("1", "10", "100", "1000"))+
  theme(legend.position = c(.8, .8))
```

```{r}
ggsave("figures/TTS_duration.png", width = 5, height = 4)
```


