---
title: "VIIRS_CalMapper_Dates"
output: 
  html_document:
    toc: TRUE
date: "2023-10-30"
---

```{r, include = F}
require(tidyverse)
require(readxl)
require(sf)
```

# Load data
This comes from `compare_VIIRS_CT_LR.Rmd`
```{r}
df <- read_rds(file.path(ref_path, 'VIIRS/CA_2020/viirs_int.rds'))
```

# Filter to CalMapper where activity end date is not the same as activity start date
```{r}
nrow_old <- nrow(df)
df <- df %>% 
  filter(ACTIVITY_END > ACTIVITY_START)
print(paste("removed", nrow_old - nrow(df), "rows"))
```

# Check dates 

## Make sure AQ_ID==199715 has dates of 7/1 to 12/31
```{r}
df %>% 
  filter(AQ_ID == 199715) %>% 
  st_drop_geometry() %>% 
  select(ACTIVITY_START, ACTIVITY_END) %>% 
  distinct()
```

## Check time zone of the date columns. This should be TRUE
```{r}
tz(df$date) == tz(df$ACTIVITY_START) 
tz(df$ACTIVITY_START) == tz(df$ACTIVITY_END)
```

# Calculate days from start of fire using `date` and `ACTIVITY_START`
```{r}
df <- df %>% 
  mutate(day_from_st = as.numeric(date - as.Date(ACTIVITY_START)))
```

## Check

Print date, ACTIVITY_START, and day_from_st
```{r}
df %>% 
  filter(AQ_ID == 199715) %>% 
  st_drop_geometry() %>% 
  select(date, ACTIVITY_START, day_from_st) %>% 
  distinct()
```

## Remove points with negative day_from_st
```{r}
nrow_old <- nrow(df)
df <- df %>% 
  filter(day_from_st >= 0)
print(paste("removed", nrow_old - nrow(df), "rows"))
```

## Remove points with date after activity end date
```{r}
nrow_old <- nrow(df)
df <- df %>% 
  filter(date < ACTIVITY_END)
print(paste("removed", nrow_old - nrow(df), "rows"))
```

# Create rank column for a nice plot
```{r}
df_rank <- df %>% 
  group_by(AQ_ID) %>% 
  arrange(duration) %>% 
  select(AQ_ID) %>% 
  distinct(AQ_ID) %>% 
  ungroup()
df_rank$rank <-  seq(1, length(unique(df$AQ_ID)))
```

```{r}
df <- left_join(df, df_rank, by = "AQ_ID")
```

```{r}
ggplot(df)+
  geom_segment(aes(x = 0, xend = duration, y = rank, yend = rank), col = "dark green")+
  geom_point(aes(x = day_from_st, y = rank), col = "red")+
  xlab("Days since start of CalMapper record")
ggsave("figures/VIIRS_CM_dates.png")
```


# Plot with normalized dates
## Create day_decimal
```{r}
df <- df %>% 
  mutate(day_decimal = (day_from_st)/as.numeric(duration))
```

## Check day_decimal 
```{r}
df %>% 
  st_drop_geometry() %>% 
  select(date, ACTIVITY_START, ACTIVITY_END, day_from_st, duration, day_decimal) %>% 
  distinct() %>% 
  head()
```


```{r}
df %>% 
  ggplot()+
  geom_histogram(aes(x = day_decimal), bins = 60, fill = "#DC3F3E")+
  xlab("Date of VIIRS detect relative to CalMappers duration")+
  theme_bw()
ggsave("figures/VIIRS_CM_dates_histogram.png", width = 9, height = 5.5)
```

## Find what the peak at around 0.8 is about
```{r}
# Find the most common day_decimal other than 0
df %>% 
  st_drop_geometry() %>%
  group_by(day_decimal) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head()
```

# Look at the AQ_IDs with day_decimal around equal to 0.810
```{r}
df %>% 
  filter(day_decimal > 0.805 & day_decimal < 0.815) %>%
  st_drop_geometry() %>%
  group_by(AQ_ID, day_decimal, date, ACTIVITY_START, ACTIVITY_END, day_from_st) %>% 
  count()
```

The answer is that there just happens to be a big fire with many VIIRS points on one day.

# Find the bar I want to map
```{r}
df %>% 
  arrange(desc(rank)) %>% 
  select(AQ_ID, ACTIVITY_NAME, ACTIVITY_START, duration, rank) %>%
  st_drop_geometry() %>% 
  distinct() %>% 
  head()
df %>% 
  filter(AQ_ID == 271821 | AQ_ID == 199715) %>% 
  st_drop_geometry() %>%
  select(date, ACTIVITY_NAME, ACTIVITY_START, ACTIVITY_END, day_from_st, duration, day_decimal) 
```

```{r}
# Add a label to each bar showing the AQ_ID
ggplot(df)+
  geom_segment(aes(x = 0, xend = duration, y = rank, yend = rank), col = "dark green")+
  geom_point(aes(x = day_from_st, y = rank), col = "red")+
  xlab("Days since start of CalMapper record")+
  geom_text(aes(x = 0, y = rank, label = AQ_ID), hjust = 0, vjust = 0)
```

