---
title: "Q2_VIIRS_wo_records"
author: "Lisa Rosenthal"
date: "2023-09-19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(fs)
library(terra)
library(mapview)
library(fasterize)

```

### Functions
```{r}

# manipulations dont like it when geometries aren't valid. and they don't work with NA validity.
make_valid <- function(sf_object) {
  validity <- st_is_valid(sf_object)
  if (sum(is.na(validity)) > 0) {
    stop("can't proceed. Contains rows where validity is NA.")
  } else{
    if (all(validity)) {
      return(sf_object)
    } else{
      sf_object <- st_make_valid(sf_object)
      print(glue::glue("there were {sum(validity == F)} not valid geometries"))
      return(sf_object)
    }
  }
}

# transform to CA crs, select fields you want, make valid
prep_polys <- function(polygon, columns, points = CA){
  message("transforming polygons")
  polygon <- polygon %>% 
  st_transform(st_crs(points)) %>% 
  select(all_of(columns))
  
  message("making polygons valid")
  make_valid(polygon)
}

```


## Goal

I want to see where VIIRS points are landing outside of recorded activity. I will want a high quality dataset of VIIRS with as few false positives as possible. 

Gather data that might be important for reducing/explaining false positives. 

For polygon layers that are wall-to-wall, convert them to rasters. For rasters in general, will want to resample at a resolution of 375 m. For categorical values, use the mode (most common pixel) when aggregating. 

Update: I tried doing that in R, but fasterize (the only one that doesn't take forever) can't calculate mode. I then tried doing it in arcpy, but can't figure out the error. So did it manually in ArcPro.


## Load CA boundary data

```{r}
# California boundary
ca_path <- file.path(ref_path, 'CA boundary/ca-state-boundary/CA_State_TIGER2016.shp')
CA <- st_read(ca_path) 

```

## raster data

layers
- responsibility areas
- non-ag data (crops + water)

Make sure the resolution is 375. RA should already be that, so convert the crop data to that. 

```{r}

# crop to CA
f_crop_to_CA <- function(raster){
  CA_t <- st_transform(CA, crs(raster))
  raster_CA <- crop(raster, CA_t)
  return(raster_CA)
}

# responsibility areas--local, state, federal
RA <- file.path(ref_path, 'Responsibility areas/California_State_Responsibility_Areas/sra_raster') %>% 
  rast
crs(RA) <- "epsg:3857" # not sure why that wasn't assigned in arcpro

# NASS crop layer
cdl_path <- file.path(ref_path, "NASS/CDL/2020_30m_cdls") %>% dir_ls(glob = '*.tif')
cdl_NASS <- rast(cdl_path)
cdl_NASS_CA <- f_crop_to_CA(cdl_NASS)

# create a mask by identifying classes to keep
levels_table <- levels(cdl_NASS_CA)[[1]]
# levels_table$Class_Names
NWL <- c('Pasture/Grass', 'Forest', "Shrubland", "Barren", "Christmas Trees",
              "Nonag/Undefined", "Deciduous Forest", "Evergreen Forest", "Mixed Forest",
              "Grassland/Pasture", "Woody Wetlands", "Herbaceous Wetlands", "Wetlands"
              )
developed <- levels_table$Class_Names[grep("devel", levels_table$Class_Names, ignore.case = T)]
other <- c("Background", "Clouds/No Data", "Water", "Nonag/Undefined", 
           "Open Water", "Perennial Ice/Snow", "Barren")
crop <- levels_table$Class_Names[!levels_table$Class_Names %in% c("", NWL, developed, other)]

# reclassify the raster to these categories
get_idx <- function(nms, newval){
  idx <- levels_table$value[levels_table$Class_Names %in% nms]
  cbind(idx, newval)
}
mat1 <- list(NWL, developed, crop, other) %>% 
  map2(., 1:length(.), get_idx) %>% 
  do.call("rbind", .)

# reclassify the raster to these categories
# 1=NWL, 2=developed, 3=crop, 4=other
cdl_CA_mask <- classify(cdl_NASS_CA, mat1, others = 4)
plot(cdl_CA_mask)


# reproject to RA's raster. output will be most common class 
cdl_CA_reproj <- project(x = cdl_CA_mask, # input
        y = RA, # the template
        method = "mode"
        )
levels(cdl_CA_reproj)[[1]] <- data.frame(
  value = 1:4, 
  Class_Names = c("nonag_veg", "developed", "crop", "other"))
names(cdl_CA_reproj) <- "CDL"
plot(cdl_CA_reproj)


# export 
writeRaster(cdl_CA_reproj, file.path(ref_path, "NASS/CDL/CDL_reclassifed_reprojected.tif"), overwrite = T)


# bind together the rasters into a brick
rasts <- c(RA, cdl_CA_reproj)
# plot(rasts)
```



## Vector based layers

Layers you might want are...
- counties (polygons)
- campgrounds (points)
- power plants (points)
- solar farms (polygons)
- landfills (polygons)

Make sure the data is valid and buffer by the VIIRS resolution divided by 2. Put into a list and export. 

```{r}
# load data, make sure it's valid and same crs as viirs, buffer by 375/2
radius <- 375/2

# counties
counties <- file.path(ref_path, "Counties") %>% 
  dir_ls(recurse = T, glob = "*.shp") %>% 
  st_read
counties_valid <- prep_polys(counties, c("NAME"))

# campgrounds
camping_FS <- file.path(ref_path, 'camping/USFS') %>% 
  dir_ls(recurse = T, glob = '*shp') %>% 
  st_read %>% 
  st_transform(st_crs(CA))
camping_CA <- file.path(ref_path, 'camping/CA_state_parks') %>% 
  dir_ls(recurse = T, glob = '*shp') %>% 
  st_read %>% 
  st_transform(st_crs(CA))
camping_FS <- filter(camping_FS, grepl('camp', PARENTACTI, ignore.case = T))
camping_FSCA <- camping_FS[CA,]
camping <- bind_rows(
  transmute(camping_FSCA, camping = 'USFS'),
  transmute(camping_CA, camping = "CA")
)
camping_valid <- prep_polys(camping, 'camping') %>% st_buffer(radius)

# power plants--not sure how to filter these. any power plants have gas flares, even solar.
# larry hunsinger
powerplants <- file.path(ref_path, 'powerplants') %>% dir_ls(glob = "*geojson") %>% st_read
# remove retired plants, make valid
powerplants_valid <- powerplants %>% 
  filter(Retired_Plant != 1) %>% 
  prep_polys(., c('PriEnergySource', 'Capacity_Latest') ) %>% 
  st_buffer(radius)

# solar plants
solar <- file.path(ref_path, 'solar')%>% dir_ls(glob = "*geojson") %>% st_read
solar_valid <- prep_polys(solar, "Combined_Class") %>% 
  st_buffer(radius)

# landfills--from https://gis.data.ca.gov/datasets/OEHHA::landfillboundaries/explore, link from Amy Klug
landfills <- file.path(ref_path, 'landfills') %>% dir_ls(glob = "*shp") %>% st_read
landfills_valid <- prep_polys(landfills, c('SiteName', "FacIT_ID")) %>% 
  st_buffer(radius)

# functions to extract data at VIIRS points ============
vector_layers <- list(counties = counties_valid, 
     camping = camping_valid,
     powerplants = powerplants_valid, 
     solar = solar_valid, 
     landfills = landfills_valid)


```

Export the data

```{r}

write_rds(vector_layers, 
          file.path('outputs_spatial/vector', 'extract_viirs.rds'))
writeRaster(rasts, 
          file.path('outputs_spatial/raster', 'extract_viirs.tif'), overwrite = T)

```



